---
title: "Minimum wages and teen employment: recovering partial homogeneity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Minimum wages and teen employment: recovering partial homogeneity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  eval = requireNamespace("did", quietly = TRUE)
)
```

This vignette works through the paper's first application: the county-level
analysis of minimum wages and teen employment from Callaway and Sant'Anna
(2021). It is the case where the cohort-time effects carry genuine,
recoverable heterogeneity, and the method finds a partially homogeneous
structure that roughly halves the variance of the pooled cells while leaving
the headline conclusion intact.

```{r libs}
library(phdid)
library(did)
data(mpdta, package = "did")
```

## The first stage

`phdid` does not estimate cohort-time effects itself. It takes them from
whichever heterogeneity-robust estimator suits the design, together with their
**joint** sampling covariance. Here that is the Callaway–Sant'Anna estimator,
whose influence functions give the exact covariance.

Fit with `bstrap = FALSE` so that the analytical covariance is computed.

```{r first-stage}
first <- att_gt(
  yname = "lemp", tname = "year", idname = "countyreal",
  gname = "first.treat", control_group = "notyettreated",
  data = mpdta, bstrap = FALSE, cband = FALSE
)

d <- ph_data(first)
d
```

`ph_data()` has kept the seven post-treatment cells, taken
$\hat\Sigma = V_{\text{analytical}} / n$, and weighted the cells by cohort size
so that the overall effect is population weighted.

The covariance is dense, and that matters. The cells share counties, so their
estimates are correlated. Remark 1 of the paper is explicit that using the
exact cross-cell covariance is what makes the reported intervals honest: the
diagonal shortcut understates the posterior variance of aggregates and can
misstate individual co-clustering probabilities badly.

## Is there anything to recover?

Ask before grouping. Both estimators will partition pure noise if asked to.

```{r homogeneity}
homogeneity_test(d)
```

The common-effect null is decisively rejected, and roughly four fifths of the
observed cross-cell dispersion is genuine rather than sampling noise. The
descriptive signal-to-noise ratio of 2.25 is the paper's observation that the
cross-cell dispersion (0.050) is more than twice the typical standard error
(0.022).

A caution the paper makes explicitly: this descriptive ratio is *not* the
separation $\delta$ of the simulation section, which is defined on the true
adjacent-group gaps and the full joint covariance. It signals that there is
recoverable heterogeneity without fixing its exact position on that scale. The
co-clustering results below assess that directly.

## One partition to report

```{r l0}
fit_l0 <- l0_ph(d)
fit_l0
```

This reproduces Table 6 of the paper. The seven cells collapse into four effect
levels: a near-zero group holding the 2004 impact effect and the recent
cohorts' effects, the 2004 cohort's larger mature effects, its peak, and the
lone positive cell.

Pooling roughly halves the variance of the cells that share a group. Notice
that even the singleton cells gain precision — in a non-orthogonal design the
Gauss–Markov estimator borrows strength across the correlated cells, which is
why Proposition 1's dominance is not confined to groups of size two or more.

```{r l0-plot}
plot(fit_l0)
```

The overall effect is essentially unchanged and slightly more precise:

```{r l0-agg}
rbind(
  flexible = aggregate(flex_twfe(d), "overall"),
  l0 = aggregate(fit_l0, "overall")
)
```

The substantive conclusion — a roughly 4% teen-employment reduction — survives,
while the estimates are summarised by a handful of interpretable levels.

## Inference that admits the partition is uncertain

`l0_ph()` commits to one grouping and its intervals condition on that grouping
being correct. With only seven correlated cells that is a strong thing to
assume. The Dirichlet Process posterior instead averages over partitions.

```{r bayes}
fit <- bayes_ph(d, alpha = 1, iters = 20000, burn = 2000, seed = 7,
                progress = FALSE)
fit
```

```{r bayes-agg}
aggregate(fit, "overall")
```

### Which groupings are firm?

```{r coclust}
round(coclustering(fit), 2)
plot(fit)
```

The small-effect cells recur together with moderate-to-high probability: the
2004 impact effect co-clusters with the lone positive 2006:2006 cell at 0.86
and with the 2004:2005 effect at 0.78, so a low-effect cluster is a robust
feature. The large mature 2004 effects are more distinct, with the peak
2004:2006 cell co-clustering with every other cell at 0.53 or below.

The matrix is more diffuse than any single partition. That is not a failure —
it is the correct report that with only seven correlated cells the finer
grouping is genuinely uncertain. This is the honest reading the co-clustering
is meant to deliver: a recurring coarse structure with calibrated uncertainty
about the finer detail.

## Checking the sampler

With seven cells there are only $B_7 = 877$ partitions, so the posterior can be
computed exactly and the sampler checked against it.

```{r enumerate}
exact <- enumerate_partitions(d, alpha = 1)
exact

data.frame(
  quantity = c("E[# groups]", "overall effect", "interval lower",
               "interval upper"),
  gibbs = c(fit$m_mean, aggregate(fit, "overall")$estimate,
            aggregate(fit, "overall")$conf.low,
            aggregate(fit, "overall")$conf.high),
  exact = c(exact$m_mean, exact$overall$mean, exact$overall$lower,
            exact$overall$upper)
)

cat("largest co-clustering discrepancy:",
    round(max(abs(fit$coclust - exact$coclust)), 4), "\n")
```

The two agree to within Monte Carlo error, which is the Appendix E benchmark.

It is worth seeing what makes that agreement possible:

```{r cov-check}
cmp <- covariance_check(d, alpha = 1, iters = 4000, burn = 1000, seed = 1)
cmp$summary
```

The diagonal shortcut reproduces the aggregates but moves individual
co-clustering probabilities, which is why the exact form is the default.

## How much does the prior matter?

With only seven cells, quite a lot for the aggregate. Rather than fix `alpha`,
report the path.

```{r alpha, fig.height = 7}
sens <- alpha_sensitivity(
  d, alpha_grid = c(0.1, 0.25, 0.5, 1, 2, 5, 10, 25, 50, 100),
  iters = 4000, burn = 1000, seed = 3
)
sens
plot_alpha_sensitivity(sens)
```

The effect slides monotonically from about −0.017 under heavy pooling toward
the flexible estimate as `alpha` grows, and is bounded away from zero only for
`alpha` above roughly 1. As a reference point, the BIC pass above selected four
groups, which the posterior expected number of groups matches at `alpha`
around 14.

That is offered as a reference, not a recipe. Choosing `alpha` from the data
turns the prior into a data-dependent object, with the usual consequences of
empirical Bayes — understated posterior uncertainty and a double use of the
data — so showing the sensitivity is preferable to committing to a single
value.

## The event-study reading

```{r dynamic}
aggregate(fit, "dynamic")
```

The effect is small at impact and accumulates over the horizon, the familiar
pattern of a minimum-wage effect that builds over several years.

One caveat on interpretation: the composition of the event-time aggregate
changes across the horizon. Event times 2 and 3 are identified off the 2004
cohort alone, so the longer-run trajectory is that cohort's, and the design
cannot say whether later cohorts would follow it. The impact-year effects, by
contrast, are small and similar across cohorts. The partial homogeneity the
method recovers here is largely the homogeneity of the *short-run* response
across cohorts, with the accumulating longer-run effects carried by the 2004
cohort's later cells kept separate.

## References

Arora, P. and Wagle, R. (2026). Partial Homogeneity in Staggered
Difference-in-Differences.

Callaway, B. and Sant'Anna, P. H. C. (2021). Difference-in-Differences with
Multiple Time Periods. *Journal of Econometrics* 225(2), 200–230.
