---
title: "When partial homogeneity helps, and when it does not"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{When partial homogeneity helps, and when it does not}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

The estimators in this package buy precision by pooling cohort-time effects
that share a common value. Whether that trade is worth making depends entirely
on whether the distinct effects are separated enough to be told apart. This
vignette reproduces, at reduced scale, the Monte Carlo of the paper's
Section 4, which is where that boundary is mapped.

```{r libs}
library(phdid)
set.seed(1)
```

## The design

```{r design}
des <- ph_design(N = 2000, T = 10, cohorts = c(0, 3, 5, 7))
des
```

Four cohorts, one never treated, giving `r 18` post-treatment cells. The unit
and time fixed effects are removed by the within transformation, so the
estimators' sampling behaviour is driven entirely by the errors.

## The two dials

The truth is parameterised by two numbers.

`m_star` is the true number of distinct effects: 1 is fully homogeneous, where
pooling is correct; `K` is fully heterogeneous, where the flexible estimator is
correct; anything in between is the partial homogeneity that motivates the
method.

`delta` is the **separation**: the distance between adjacent group means in
standard errors of the flexible estimates. It is unit-free, and it is the
variable that decides whether the partition can be recovered at all.

```{r truth}
truth <- ph_truth(des, m_star = 6, delta = 6)
table(truth$labels)
round(sort(unique(truth$tau)), 3)
```

A single draw is a `ph_data` object like any other, so every estimator in the
package applies to it:

```{r draw}
d <- ph_sample(des, truth)
l0_ph(d)$m          # groups recovered
adjusted_rand(l0_ph(d)$partition, truth$labels)
```

## Precision under partial homogeneity

The headline claim is that at partial homogeneity the feasible estimators cut
the sampling variance of the cohort-time effects substantially relative to
flexible TWFE while remaining essentially unbiased.

The Bayesian estimator dominates the run time, since each replication runs a
full chain. The paper uses 500 replications; here we use far fewer so the
vignette builds quickly. Expect noticeably noisier numbers than the published
tables.

```{r sim-main}
res <- sim_study(
  des, m_star = 6, delta = 6, R = 25,
  bayes_args = list(iters = 400, burn = 100),
  seed = 11, progress = FALSE
)
res[, c("method", "var_ratio", "abs_bias", "ari", "cover_CATT", "cover_ATT")]
```

Three patterns to look for, all of which the full-scale experiment shows
sharply.

**Pooled TWFE is always the most precise and always unusable for cohort-time
effects.** Its variance ratio sits near 0.14 regardless of the truth, but under
any heterogeneity it is severely biased, and its intervals -- short and centred
on the wrong estimand -- cover at around 0.04.

**At partial homogeneity the feasible estimators approach the infeasible
oracle.** In the paper's full experiment they cut the CATT sampling variance by
26-52% relative to flexible TWFE while remaining essentially unbiased, and
recover the partition well.

**They degrade correctly at the extremes.** At `m_star = 1` they shrink toward
pooling; at `m_star = K` any pooling hurts and flexible is preferred.

## The separation boundary

This is the part that matters most in practice. Fix the number of distinct
effects and vary how far apart they are.

```{r sim-delta}
grid <- do.call(rbind, lapply(c(3, 6, 12), function(delta) {
  sim_study(des, m_star = 6, delta = delta, R = 15,
            methods = c("flexible", "oracle", "l0"),
            seed = 100 + delta, progress = FALSE)
}))
grid[, c("method", "delta", "var_ratio", "ari")]
```

At `delta = 3` the partition cannot be recovered reliably, and the selection
noise erases -- indeed reverses -- the variance advantage: the feasible
estimators are *less* precise than flexible TWFE. As separation grows the gap
to the oracle closes, and by `delta = 12` the l0 estimator essentially attains
the oracle variance reduction with near-perfect recovery.

The practical message is that these methods deliver their promised gains when
heterogeneity is "clumpy" -- a modest number of well-separated effect levels --
and should not be expected to when the effects form a near-continuum.

## Where the two estimators genuinely differ

Not on variance, but on inference. When the partition is uncertain, the l0
plug-in interval treats a possibly-wrong grouping as known, and its coverage
collapses; the posterior averages over partitions and degrades gracefully.

```{r sim-coverage}
cov_grid <- do.call(rbind, lapply(c(3, 12), function(delta) {
  sim_study(des, m_star = 6, delta = delta, R = 15,
            methods = c("flexible", "l0", "bayes"),
            bayes_args = list(iters = 400, burn = 100),
            seed = 200 + delta, progress = FALSE)
}))
cov_grid[, c("method", "delta", "cover_CATT", "len_CATT", "cover_ATT")]
```

In the paper's full experiment, at low separation the l0 plug-in CATT coverage
falls to 0.57-0.62 while the Bayes-PH credible intervals hold 0.79-0.81; where
the effects are separable both attain near-nominal coverage.

Two further observations from that experiment are worth carrying into applied
work. The **overall ATT is well calibrated everywhere** for both estimators,
because aggregation averages out the cell-level errors -- it is the cohort-time
intervals where the regime bites. And honest uncertainty under partition
selection comes not from adjusting the variance but from averaging over the
partition, which is what the posterior does and what no standard-error
correction to the l0 interval can replicate. Sample splitting does not rescue
it either: halving the sample degrades recovery, and a wrongly merged partition
injects a specification bias that a correctly sized standard error cannot undo.

## Reproducing a published row

The tables in the paper use 500 replications and the full Gibbs settings. One
row costs hours rather than seconds:

```{r full, eval = FALSE}
res_full <- sim_study(
  des, m_star = 6, delta = 6, R = 500,
  bayes_args = list(iters = 4000, burn = 1000),
  seed = 1234
)
```

## References

Arora, P. and Wagle, R. (2026). Partial Homogeneity in Staggered
Difference-in-Differences, Section 4.
